Mistriage

### 2019-05-28 ManuscriptMarkdown.Rmd
Add files via upload

---
title: "Mistriage"
author: "Martin Henriksson"
date: "190517"
output: word_document
---

@@

Trauma care involves many levels of the health care system. In a typical, high resource setting, the initial management of patients is performed on the scene of the trauma by Emergency Medical Services (EMS). In this prehospital phase; patient data, vital signs and scene information is collected and transferred to the receiving hospital. This information is then evaluated with the help of a system to determine the level of trauma and prepare adequate resources (4). The level of trauma dictates if a full or limited trauma team is activated (5).

Systems to determine the level of trauma can be based on clinical prediction models. Such models are continuously being developed and studied, examining different parameters and different contexts concerning trauma care. These models differ in quality and characteristics, but they generally perform well when predicting survival (6). Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then supposed to be implemented in different settings, such as non-trauma centres.

What is not fully understood is how this transfer affects the performance of these models. Previous research that studied the transfer of clinical prediction models between settings has shown that model performance in terms of calibration can be adversely affected (7). However, the aforementioned research studied model transfers between substantially different settings and did not assess more clinically relevant performance measures, such as misclassification.

In trauma, misclassification is often referred to as mistriage. Here, triage refers to the classification of severity as minor or major trauma. Mistriage can be subdivided into overtriage; the incorrect classification of a patient with minor trauma as major trauma, or undertriage; the incorrect classification of a patient with major trauma as minor trauma. Both are detrimental in terms of patient care and distribution of resources.

@@

# Materials and methods
## Study design
We conducted a registry-based cohort study. Sweden has a nationally encompassing register for registering trauma called SweTrau. We used SweTrau data to create clinical prediction models, which we then transferred between different contexts to study the effects on mistriage. Please refer to appendix I for a brief explanatory list of the statistical terminology.

@@

The trauma care organization in Sweden is usually clearly defined in major metropolitan areas (albeit the specific routines differ), and trauma patients are transported directly to a predesignated hospital with specific trauma competency and capacity. In large, more rural parts of the country, trauma patients are instead transported to the nearest hospital for stabilization. Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity (19). Most hospitals in Sweden use some form of system to categorize the level of trauma and necessary response. These systems are usually based on the patient's vital signs and injury mechanism, as reported by the EMS before patient arrival, or as registered in the emergency department if the patient arrives by other means (16).

We used data from SweTrau between 2011 and 2016, as this study’s inception was in 2016 and data obtained the same year. Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau. Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives (20). Included hospitals are both university and non-university hospitals, all of which receive trauma patients.

Patient data is recorded upon arrival to the hospital. All hospital personnel can perform patient registration, but it is recommended they received Abbreviated Injury Scale (AIS) training. The SweTrau inclusion criteria are: Traumatic event with subsequent activation of hospital trauma protocol, admitted patients with NISS (New Injury Severity Score) > 15 and patients transferred to the hospital within 7 days of traumatic event with NISS > 15. SweTrau excludes patients if the only traumatic event is chronic subdural hematoma or if hospital trauma protocol is activated without traumatic event (21).

## Participants 
The eligibility criteria are patients registered in SweTrau, and age 15 or above. This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care (22). We recognize this age cut-off is not completely without controversy. However, in Sweden patients age ≥15 go to the adult emergency rooms and the Swedish guidelines for trauma activation define children as age <15 (16). Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off (9). Patient age was obtained from the SweTrau register.

## Variables
### Model predictors
Our clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital. The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score (23).

### Model outcome
The outcome that we used to develop the clinical prediction models was all cause mortality within 30 days of the trauma.

### Participant characteristics
To describe the patient cohort we reported age, sex, ASA physical status classification system, Injury Severity Score (ISS) and New ISS (NISS).

### Study outcome
We used ISS > 15 as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma (24). We defined overtriage as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma. Clinically, in the event of overtriage, a full trauma team is activated and tasked with the care of a minimally injured patient. The opposite is true with undertriage, where a patient with serious injuries is managed with (potentially) inadequate resources  (10). We defined the overtriage rate as the number of overtriaged patients divided by all patients. We defined the undertriage rate as the number of undertriaged patients divided by all patients. The mistriage rate was defined as the sum of the over- and undertriage rates.

## Data sources and measurements
Model predictors, outcome, participant characteristics and study outcome as outlined above will all be obtained from the SweTrau registry. The method of measurement for the model predictors is not specified (for example, if SBP is measured using an automated cuff or manually) in the registry entries. In Swedish emergency rooms, patient parameters are usually obtained by a registered nurse or assistant nurse and are assumed to be accurate. 
Whether the patient is dead or alive 30 days after trauma is manually entered into the registry locally, by each respective hospital. Foreign citizens discharged within 30 days are registered as survivors. Participant characteristics; sex and age are obtained from the patient file before being registered in SweTrau. ASA, NISS and ISS are calculated by hospital personal based on the patient’s injuries, and registered.

## Bias
Data analysis was conducted in a step-by-step fashion, and according to a prearranged analysis plan. The analysis plan and statistical analysis code was finalised using simulated data instead of the real data. These efforts were taken to avoid confirmation bias when conducting the data analysis. The analysis plan was reviewed by an experienced statistician and programmer prior to implementation to ensure objectivity. Neither outcome nor variable were blinded when conducting data analysis, which made a structured objective approach even more important.

## Study size
All patients matching eligibility criteria are listed above. Four data sets were used to study the transfer of clinical prediction models, each data set representing a different setting. Data set, and sample size considerations are outlined below.

## Quantitative variables
SBP and RR was modelled using restricted cubic splines with four knots whenever possible, placed at equally spaced percentiles and GCS as a continuous linear term. When describing the participant characteristics all quantitative variables were to be presented as continuous. ISS was also presented as dichotomous using ISS > 15 as the cutoff.

@@

### Indicidual centres
This data set was also be split into multiple samples. Each centre with large enough sample size to develop and validate a model constituted its own sample. This data set was removed in the final version of the study due to a lack of centres with a large enough sample size.

@@

The minimum sample size of development and validation samples was 140 and 200 respectively. The total number of events and non-events per data set was therefore at least 680. We performed analyses only on data sets for which all samples included at least the minimum number of patients.

### Sequence of analysis
The programming language R was used for all analyses (27), and all progress was uploaded to a GitHub repository for inspection (28). We performed the analyses in the sequence of model development, model validation and finally model comparison. These steps were repeated in each data set. Below we used the transfer of a model from a high-volume sample to a low volume sample as an example to describe the complete procedure.

### Model development
In the model development step a clinical prediction model was developed in the high-volume centre development sample. The model was developed using logistic regression as implemented in the R function glm. The dependent variable was all cause mortality within 30 days of trauma and independent variables SBP, RR, and GCS modelled as previously described. To avoid overfitting the model we used a bootstrap procedure to estimate a linear shrinkage factor that we then applied to the model coefficients (29). The shrunk model was then used to estimate the probability of all cause 30-day mortality in the development sample. We then performed a gridsearch across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% (11). This cut-off was then used to classify patients as major or minor trauma.

@@

### Model comparison
Finally, in the model comparison step, the difference in model performance between the high and low volume centre validation samples was calculated. We used an empirical bootstrap to estimate 95% confidence intervals (CI) around performance and differences in performance estimates. Both bootstrap procedures used 1000 bootstrap samples drawn with replacement of the same size as the original samples. The three steps of model development, model validation, and model comparison were repeated in all four sets of data.

@@

## Ethical considerations
This study was conducted in accordance with the four major principles of medical ethics and The Declaration of Helsinki (31, 32).

### Respect for autonomy
The respect for autonomy was upheld as Swedish Privacy Law requires the informed consent of all participants. Patients included in SweTrau receive letters detailing that data obtained may be used for scientific purposes. They are also informed that participation is voluntary and that they may withdraw at any time.

### The Principle of Beneficence
Throughout the entire study we attempted to act in the best interest of the patients. In the long run we hope this research will improve the care of trauma patients, and therefore prove beneficial to those affected. We also believe that the patients who consented to research want to contribute to this improvement of care, and as thus by performing this study we hope to honour their wishes.

### The Principle of Nonmaleficence
All attempts were made not to harm any patients included in this study, either intentionally or unintentionally.  Data leakage and patient identification constitutes the primary risk. All data obtain was depersonalized and handled with care to minimize risk of unauthorised access and subsequent patient identification. By implementing these actions, we determined that the risk to the patient population is minimal.

### The Principle of Justice
As this study contained no intervention, the principle of justice does not truly apply. However, all cases included from the registry were treated as equal. It is also our wish that any scientific gains this study may contribute will be implemented in a fair and equal fashion.

### Ethical permit
This study is approved by the regional ethics review board in Stockholm, Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32.

@@

```{r, echo = FALSE, results = "asis"} 
## Table 1 = Sample characteristics table
#
```

## Development and validation
During model development, the number of imputations used for each data set was `r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and low volume data set, the metropolitan and non-metropolitan data set and the multi and single centre data set respectively. This number was equal to the maximum percentage of missing observations for each sample in the respective data sets. Restricted cubic splines with four knots placed at equally spaced percentiles were implemented when modelling systolic blood pressure and respiratory rate. 1000 bootstrap samples drawn with replacement was used to estimate a linear shrinkage factor that was applied to the developed models to avoid model overfitting.

Table 2 shows model validation performance. Model validation performance refers to the model performance when applied in the validation subsample from the sample in which it was developed. The sample with the best validation model performance was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median`. Performance in terms of under and overtriage for the same model was `r best.validation$Validation.undertriage.median`, and `r best.validation$Validation.overtriage.median`. Model validation performance varied across all samples with a median mistriage rate range of `r min(results$original.stats$Validation.mistriage.median)` - `r max(results$original.stats$Validation.mistriage.median)`. Uncertainty interval(UI) for validation mistriage, undertriage and overtriage ranged from ... to ... across all samples.

```{r, echo = FALSE, results = "asis"} 
## Table 2 = Cropped to only show validation
table.2
```

## Comparison
Model performance after transfer was determined for each model being transferred to the other sample in the data set in all data sets (Table 3). When transferred, the model with the best performance i.e. lowest mistriage rate was the `r best.transfer$Sample.name` model with a median mistriage of `r best.transfer$Transfer.mistriage.median`. Concerning the UI for transfer mistriage, undertriage and overtriage, this too ranged from 0 to 0.830 across all samples.

```{r, echo = FALSE, results = "asis"} 
## Table 3 = Cropped to only show transfer
table.3
```

The main purpose of this study was to examine the transferability of clinical prediction models for trauma. Table 4 shows the transferred model performance when compared to the validation performance in the same sample, transferred model performance varied between samples and data sets.

The lowest model performance when compared to the sample validation performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median`, meaning this model performed worse when transferred to the other sample in the same data set, when compared with the model developed and validated in this sample. In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` more patients would be wrongly classified as major or minor trauma. However, when uncertainty intervals (UI) were obtained no observed differences between transferred model performances and validation model performances in any sample were found to be significant.

```{r, echo = FALSE, results = "asis"} 
## Table 4 = Cropped to only show transfer differance
table.4
```

# Discussion (Needs update with code snippets) 
This study aimed to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates. The most notable effects on model performance following model transfer compared to the original sample model was seen when the metropolitan model was transferred to the non-metropolitan sample. The performance of the metropolitan model transferred to the non-metropolitan sample was a mistriage rate of ..., and the validation mistriage rate of the non-metropolitan model was .... This indicates that the mistriage increased by ... when using the metropolitan model in a non-metropolitan setting. Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer. The UI of this difference however ranged from ... to ..., indicating that our findings are compatible with a ... decrease to ... increase in mistriage. This means that with the negative UI (...), the effect of transfer would be an improvement in model performance. The upper limit of the UI (...) would mean a marked decrease in model performance, especially compared to the point estimate of ....

The transfer of clinical prediction models in trauma care has not previously been extensively studied. In 2016, Gerdin et al. also found model transfer to adversely affect model performance (18). They found the transferred model to perform poorly in terms of calibration, and that this could be improved by updating the model. In our current study we did not examine model performance in terms of calibration, neither did we examine the effects of updating the model. What the two studies have in common is an observed decline in model performance, when transferring the models between certain settings. Therefore, we believe model transfer can potentially lead to a decline in model performance compared to the setting in which it was originally developed.  Model transfer should be performed with caution, and the effects examined before implementing the transferred model in a clinical setting.

External validation of clinical prediction models, meaning the evaluation of a clinical prediction model in a setting in which it was not originally developed has however been studied more extensively. The concepts of external validation are applicable and comparable to our current study, with the exception being that out models were technically externally validated using a validation subsample before being transferred to an entirely different sample. Studies using simulated clinical prediction models based on different predictors often show a decline in model performance when being externally validated (33, 34), both underlining the necessity of testing models in other populations before clinical use. A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) (35), showing models significantly underestimating the probability of CAD. Our study did not show the same dramatic decrease in model performance after transfer but combined with the results of other studies of external validation it further warrants caution when transferring models between different settings.

## Strengths and limitations
This study examined model transfer in a limited setting (Sweden, as a single health care setting), which has advantages and disadvantages. The study design realistically reflects potential transfers: clinical prediction models for trauma are usually developed in large metropolitan centres and then transferred as clinical recommendations to minor non-metropolitan centres in which traumatic injuries are not as common.

In terms of disadvantages these results are restricted to a single health care system in Sweden, and the generalizability to other health care systems might be limited. Also, the small number of patients (27043) weakens the applicability of the results on a more general setting. We originally intended to study all single centres with a valid number of events, but when performing the analysis only one centre provided a sufficient number of events for study. Leading to the loss of the fourth data set: Individual centres, and the comparison between these. We recognize that the results this data set might have provided could reflect realistic situations as major trauma centres regularly exchange data to improve patient care.

In Sweden major trauma is relatively uncommon, especially in non-metropolitan and minor centres, and therefore it is reasonable to assume this leads to inconsistencies when performing SweTrau patient registration. Such as using NISS instead of ISS, registering SBP by category and not by absolute value, not following up on patient survival, or other local differences. This might lead to exaggerated performance differences between the different settings, as the transferred models are based on registry data.

We developed models using logistic regression based on registry data, instead of using pre-developed models such as the Revised Trauma Score, eliminating the risk of potentially misrepresenting already established models. The models used in this study used GCS, RR and SBP as predictors for 30-day mortality. Potentially, the use of additional predictors such as age, mechanism of injury, or newer predictors as chock index might increase model performance. Furthermore, we used 30-day mortality as an outcome. We recognize that late mortality, time in hospital, functional impairment, and morbidity are also important outcome factors in the context of trauma care.

Statistical analysis and comparisons were performed using R. We aimed to follow a predetermined analysis plan in a step-by-step fashion with to aim of avoiding confirmation bias. However, programming challenges and late-time code inclusions required a flexible approach to this analysis plan, with revisions of previous steps. We recognize that this is not optimal but aimed to maintain objectivity by not deviating in any major way from the analysis plan. Using GitHub, all code was made public for inspection and to ensure reproducibility (28).

## Clinical applications
Clinical prediction models are constantly being developed in different contexts, and often transferred to settings in which they were not originally developed. The results of this study show that mistriage can potentially increase following model transfer, in this case more prominently when transferring from a metropolitan setting to a non-metropolitan setting. This supports the view that model transfers should be done with caution, as performance might decline.

Recognizing these risks and updating the models accordingly will lead to an increased accuracy during the triage of trauma patients. This could potentially save lives by minimizing undertriage, and likely save resources by minimizing overtriage. Potentially, because of registries such as SweTrau, analysis of model transfer could be performed continuously to optimize models currently in use and advise which model transfers might be detrimental (or potentially beneficial).

## Future studies
To increase the generalisability, studies of this type could be performed in different settings e.g. assessing the transferability between different countries, such as Sweden and the United Kingdom. Also, increasing the number of patients would also increase the power of the study results, one potential way of doing this would be to replicate this study using the very latest data in the SweTrau registry, as many hospitals have joined recently which in turn has led to an increase in the number of patients registered.

@@

In terms of equity, patient sex and age were obtained to characterize the patient cohort: 66 % male, and a median age of 41 (mean 44). No actions were undertaken in response to this division. Further studies might aim to examine if model performance differs using sex or age as a predictor.

## Conclusion (Needs update with code snippets) 
A clinical prediction model developed using logistic regression and GCS, RR and SBP as predictors can be used to predict 30-day mortality in different settings. Model transfer in the metropolitan and non-metropolitan data set lead to ... mistriage rate (95% UI ...). This was mainly due to .... In all samples, 95% UI for model transfer comparison included 0. Poor model transfer can potentially lead to increased patient mortality and a misappropriation of resources. We recognize the limitations of this study, but believe further studies are warranted due to the potentially severe patient consequences.

>>>>

---
title: "Mistriage"
author: "Martin Henriksson"
date: "190429"
output: word_document
---

@@

In a typical, high resource setting, the initial management of patients is performed on the scene of the trauma by Emergency Medical Services (EMS). Patient data, vital signs and scene information is transferred to the receiving hospital. This information is evaluated using a system to determine the level of trauma and prepare adequate resources (4), dictating if a full or limited trauma team is activated (5).

Systems to determine the level of trauma can be based on clinical prediction models. Such models are continuously being developed and studied, examining different parameters and different contexts. Models differ in quality and characteristics, but generally perform well predicting survival (6). Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then implemented in different settings, such as non-trauma centres.

What is not fully understood is how this transfer affects model performance. Previous research has shown that model performance in terms of calibration can be adversely affected (7). However, the aforementioned research studied model transfers between substantially different settings and did not assess more clinically relevant performance measures, such as misclassification.

In trauma, misclassification is often referred to as mistriage. Triage refers to the classification of severity as minor or major trauma. Mistriage can be subdivided into overtriage; incorrect classification of a patient with minor trauma as major trauma, or undertriage; incorrect classification of a patient with major trauma as minor trauma. Both are detrimental in terms of patient care and distribution of resources.

@@

# Materials and methods
## Study design
We conducted a registry-based cohort study. Sweden has a nationally encompassing register for registering trauma called SweTrau. SweTrau data was used to create clinical prediction models, which were transferred between different contexts to study the effects on mistriage.

@@

The trauma care organization in Sweden is usually clearly defined in major metropolitan areas: trauma patients are transported directly to a predesignated hospital with specific trauma competency. In large, more rural parts of the country, trauma patients are instead transported to the nearest hospital for stabilization. Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity (19). Most hospitals in Sweden use some form of system to categorize the level of trauma. These systems are usually based on patient vital signs and injury mechanism, as reported by the EMS, or as registered in the emergency department (16).

Data was obtained from SweTrau between 2011 and 2016. Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau. Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives (20).

The SweTrau inclusion criteria are: Traumatic event with subsequent activation of hospital trauma protocol, admitted patients with NISS (New Injury Severity Score) > 15 and patients transferred to the hospital within 7 days of traumatic event with NISS > 15. SweTrau excludes patients if the only traumatic event is chronic subdural hematoma or if hospital trauma protocol is activated without traumatic event (21).

## Participants 
The eligibility criteria were patients registered in SweTrau, age 15 or above. This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care (22). The Swedish guidelines for trauma activation define children as age <15 (16). Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off (9). Patient age was obtained from the SweTrau register.

## Variables
### Model predictors
The clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital. The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score (23).

### Model outcome
The outcome used to develop the clinical prediction models was all cause mortality within 30 days of the trauma.

### Participant characteristics
To describe the patient cohort we reported age, sex, Injury Severity Score (ISS) and New ISS (NISS).

### Study outcome
ISS > 15 was used as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma (24). Overtriage was defined as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma. Overtriage rate was defined as the number of overtriaged patients divided by all patients. Undertriage rate was defined as the number of undertriaged patients divided by all patients. Mistriage rate was defined as the sum of the over- and undertriage rates.

## Data sources and measurements
Model predictors, outcome, participant characteristics and study outcome as outlined above were all obtained from SweTrau. The method of measurement for the model predictors is not specified (for example, if SBP is measured using an automated cuff or manually) in the registry entries. In Swedish emergency rooms, patient parameters are usually obtained by a registered nurse or assistant nurse and are assumed to be accurate. 
Whether the patient is dead or alive 30 days after trauma is manually entered into the registry locally, by each respective hospital. Participant characteristics are obtained from the patient file before being registered in SweTrau. NISS and ISS are calculated by hospital personal based on the patient’s injuries.

## Bias
Data analysis was conducted according to a prearranged analysis plan. The analysis plan and statistical analysis code was finalised using simulated data. These efforts were taken to avoid confirmation bias. The analysis plan was reviewed by an experienced statistician and programmer prior to implementation, to ensure objectivity. Neither outcome nor variable were blinded when conducting data analysis, which made a structured objective approach important.

## Study size
All patients matching eligibility criteria listed above. Four data sets were used to study the transfer of clinical prediction models, each data set representing a different setting. Data set, and sample size considerations are outlined below.

## Quantitative variables
SBP and RR was modelled using restricted cubic splines with four knots whenever possible placed at equally spaced percentiles, and GCS as a continuous linear term. When describing the participant characteristics quantitative variables were presented as continuous. ISS was presented as dichotomous using ISS > 15 as the cutoff.

@@

### Individual centres
This data set was also split into multiple samples. Each centre with large enough sample size to develop and validate a model constituted its own sample. This data set was removed in the final version of the study due to a lack of centres with a large enough sample size.

@@

The minimum sample size of development and validation samples was 140 and 200 respectively. We performed analyses only on data sets for which all samples included at least the minimum number of patients.

### Sequence of analysis
The programming language R was used for all analyses (27), all progress was uploaded to a GitHub repository (28). Analyses were performed in the sequence of model development, model validation and finally model comparison. These steps were repeated in each data set. Below we used the transfer of a model from a high-volume sample to a low volume sample as an example to describe the complete procedure.

### Model development
In the model development step a clinical prediction model was developed in the high-volume centre development sample. The model was developed using logistic regression as implemented in the R function glm. The dependent variable was all cause mortality within 30 days of trauma and independent variables SBP, RR, and GCS modelled as previously described. To avoid overfitting the model we used a bootstrap procedure to estimate a linear shrinkage factor that we then applied to the model coefficients (29). The shrunk model was then used to estimate the probability of all cause 30-day mortality in the development sample. A gridsearch was performed across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% (11). This cut-off was then used to classify patients as major or minor trauma.

@@

### Model comparison
Finally, in the model comparison step, the difference in model performance between the high and low volume centre validation samples was calculated. Empirical bootstrap was used to estimate 95% confidence intervals (CI) around performance and differences in performance estimates. Both bootstrap procedures used 1000 bootstrap samples drawn with replacement of the same size as the original samples. The three steps of model development, model validation, and model comparison were repeated in all four sets of data.

@@

## Ethical considerations
This study is approved by the regional ethics review board in Stockholm, Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32. All patients included recived letters detainling SweTrau inclusion, and the possibility of their data being used for scientific purposes. It was assumed that the patients who consented to research want to contribute to this improvement of care, and as thus by performing this study the authors hope to honour their wishes.

@@

```{r, echo = FALSE, results = "asis"} 
## Table 1 = Sample characteristics table
##result.tables
```

## Development and validation
During model development, the number of imputations used for each data set was `r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and low volume data set, the metropolitan and non-metropolitan data set and the multi and single centre data set respectively.

Table 2 shows model validation performance. The sample with the lowest mistriage was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median`. Performance in terms of under and overtriage for the same model was `r best.validation$Validation.undertriage.median`, and `r best.validation$Validation.overtriage.median` respectivly. Across all samples model validation performance in terms of mistriage ranged from `r min(results$original.stats$Validation.mistriage.median)` to `r max(results$original.stats$Validation.mistriage.median)`.

**Table 2** Model validation performance  
```{r, echo = FALSE, results = "asis"} 
## Table 2 = Cropped to only show validation
table.2
```

## Comparison
Model performance after transfer was determined for each model being transferred to the other sample in the data set in all data sets (Table 3). When transferred, the model with the lowest mistriage rate was the `r best.transfer$Sample.name` model with a median mistriage of `r best.transfer$Transfer.mistriage.median`.

**Table** 3 Model transfer performance  
```{r, echo = FALSE, results = "asis"} 
## Table 3 = Cropped to only show transfer
table.3
```

Table 4 shows the transferred model performance when compared to the validation performance in the sample reciving the transferred model, e.g. high volume transfer performance minus low volume validation performance.

The highest mistriage when compared to the reciveing sample validation model performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median`, meaning this model performed worse when transferred to the other sample in the same data set.. In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` more patients would be wrongly classified as major or minor trauma.

**Table 4** Comparison mistriage (Sample validation performance subtracted from transferred performance)  
```{r, echo = FALSE, results = "asis"} 
## Table 4 = Cropped to only show transfer differance
table.4
```

# Discussion (Needs update with code snippets) 
This study aimed to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates. The most notable effects on model performance following model transfer were seen after transferring the metropolitan model. Mistriage increased by ... when using the metropolitan model in a non-metropolitan setting. Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer. For this particular transfer 95% UI ranged from ... to ..., indicating that our findings are compatible with a -0.055 decrease to 0.630 increase in mistriage. The negative 95% UI of ... would signifiy an improvement in model performance. The higher 95% UI of ... would mean a marked increase in model mistriage, compared to the point estimate of ....

The transfer of clinical prediction models in trauma care has not previously been extensively studied. In 2016, Gerdin et al. found model transfer to adversely affect model performance in terms of calibration, but that this could be improved by updating the model (18). This study did not examine model performance in terms of calibration, neither was the effects of updating the model explored.

External validation as the evaluation of a clinical prediction model in a setting in which it was not originally developed has been studied more extensively. Studies using simulated clinical prediction models based on different predictors showed a decline in model performance when being externally validated (33, 34). A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) (35), showed models significantly underestimating the probability of CAD. This study did not find the same dramatic decrease in model performance after transfer. However, the combined results of this study, the studies on model transfer and the studies on external validation warrants further caution when transferring models between different settings.

## Strengths and limitations
This study design realistically reflects potential transfers: e.g. clinical prediction models for trauma are usually developed in large metropolitan centres and then transferred as clinical recommendations to minor non-metropolitan centres in which traumatic injuries are not as common.

Originally all single centres with a valid number of events were to constitute one data set, but when performing the analysis only one centre provided a sufficient number of events. Leading to the loss of the fourth data set: Individual centres, and the comparison between these. Possibly, the results this data set might have provided results that could reflect realistic situations as major trauma centres regularly exchange data to improve patient care.

In Sweden major trauma is relatively uncommon, this might lead to inconsistencies across centres when performing SweTrau registration. E.g. using NISS instead of ISS, registering SBP by category and not by absolute value or not following up on patient survival. This might lead to exaggerated performance differences between the settings.

The models in this study used GCS, RR and SBP as predictors for 30-day mortality. Potentially, the use of additional predictors such as age, mechanism of injury, or newer predictors as shock index might increase model performance. Furthermore, we used 30-day mortality as an outcome, recognizing that late mortality, time in hospital, functional impairment, and morbidity are also important outcome factors in the context of trauma care.

The predetermined analysis plan was to be followed in a step-by-step fashion, to avoiding confirmation bias. Programming challenges and late-time code inclusions required a flexible approach to this analysis plan, sometimes with revisions of previous steps. This is not optimal, and efforts were made to not deviate in any major way. Using GitHub, all code was made public for inspection and to ensure reproducibility (28).

## Clinical applications
The results of this study show that mistriage can potentially increase following model transfer, in this case more prominently when transferring from a metropolitan setting to a non-metropolitan setting. This supports the view that model transfers should be done with caution, as performance might decline.

Recognizing these risks and updating models accordingly will lead to increased accuracy during trauma patient triage. Potentially saving lives by minimizing undertriage, and likely save resources by minimizing overtriage. With registries such as SweTrau, analysis of model transfer could be performed continuously to optimize models currently in use and advise which model transfers might be detrimental (or potentially beneficial).

## Future studies
Studies on model transfer can be performed in different settings, e.g. assessing the transferability between different but similar countries, such as Sweden and Norway. Also, increasing the number of patients would increase the power of the results, one way of doing this would be to replicate this study with the latest data in the SweTrau, as many hospitals have joined recently leading to an increased number of patients registered.

@@

## Conclusion (Needs update with code snippets) 
Model transfer from the metropolitan sample to the non-metropolitan sample led to an increased mistriage rate of ... (95% UI ...). Mainly due to a marked increase in overtriage. Across all samples, 95% UI for model transfer comparison included .... Model transfer can potentially lead to increased patient mortality and a misappropriation of resources, due to increased undertriage and overtriage respectively. Recognizing the limitations of this study, the authors believe further studies are warranted due to the potential economic costs and patient consequences of poor model transfers.
